$ _{} $ $ _{} $ $ _{} $ $ _{} $ $ _{} $In the digital age, text is one of the main means of communication and especially of transmitting knowledge. In 2018, it was about $ 80\% $ of the information circulating on the web. But time being a scarce commodity, one would like to have the possibility to access directly the salient information of texts, or just to have an overview of it before spending time on it. Hence the need for a powerful automatic summarization system.
In this work, our objective is to set up an efficient automatic summarization system. For this purpose, we have considered both abstractive and extractive summarization, leaving the choice to the user, according to his information needs.
For extractive summarization, we considered two approaches which consisted in using a python module called \textit{gensim} for synthesis, and then a mixture of several synthesis algorithms which we called \textit{merging}. For the abstract summarisation, we opted for the \textit{transformer} type models as they have so far achieved the highest performance for various natural language processing tasks. Nevertheless, we have served the \textit{transformers} used for abstractive summarization through a pipeline to improve performance and increase the number of receivable pages.
In our experiments, we found that the \textit{merging} algorithm performed qualitatively better than \textit{gensim}. We also recorded a clear improvement in the results provided by the \textit{transformers} when served through the pipeline we proposed. Thus, we obtained a system capable of rendering abstractive or extractive summaries of texts. For this purpose, we have set up an application programming interface that developers could use to perform automatic summarization in their respective systems, we have implemented a web application that exploits the services of this developed interface and we have made a model that will be improved as we collect text-summary pairs through our system.\\
$ _{ } $\\
\underline{\textbf{Keywords}} : Artificial intelligence, Encoder-decoder, Transformer, automatic text sum\-ma\-ri\-za\-tion, Python, Flask, JavaScript, ReactJS, ExpressJS.